MetaTracts - A Method for Robust Extraction and Visualization of Carbon Fiber Bundles in Fiber Reinforced Composites

### 2014-09-22 chapters/MetaTracts.tex
Local old commit
Updated text
local text commit

\begin{figure}[tb]
%\includegraphics[width=0.45\textwidth]{imagesMT2014/algo.png}
\includegraphics[width=0.5\textwidth,  trim = 0mm 300mm 0mm 0mm, clip,]{imagesMT2014/image_algo}
\caption{Flow chart of the MetaTracts approach for fiber bundle extraction}
\label{fig:flowchart}
\end{figure}

@@

\subsection {MetaTracts generation}
\label{subsec:MetaTracts-generation}
In $\mathbb{R}^2$ all the above properties hold except that cylinders are replaced by rectangles. We explain the process of MetaTracts generation in $\mathbb{R}^2$. The procedure trivially extends to $\mathbb{R}^3$. The reliable Hessian step  generates a local orientation in $\mathbb{R}^2$ and a $R_H$[0,1]. In Figure \ref{fig:algo} all regions which have reliable Hessian less than the threshold have been marked in blue(Fig.~\ref{fig:algo}B). Let the seed point associated with the MetaTract be grid point $C_{p}$ (Property 2, Section~\ref{subsec:mtprop}). The local orientation at $C_p$ as computed in the step above is $N_P$ and is given by the dark green arrow (Fig.~\ref{fig:algo}C). We generate a rectangle of length L and radius R, the green dashed rectangle shows the rectangle generated at $C_p$ (Property 3 and 5, Section~\ref{subsec:mtprop}). 
\begin{figure}[tb]
  \begin{center}
   \includegraphics[ width=0.45\textwidth]{imagesMT2014/image4}
 % \includegraphics[ width=0.45\textwidth]{imagesMT2014/MT_abstract} 
  \end{center}
  \vspace{-1em}
 \caption{MetaTracts generation: (B) MetaTract generation, region with $R_H$ less than threshold. (A) close up of marked region in (B). (C) Parts of a single MetaTract.}
  \label{fig:algo}
    \vspace{-1.5em}
\end{figure}
The set of vertices which are in the green region but not overlapped by blue are possible candidates for the start point of the next cylinder in a particular MetaTract. We call these vertices \textit{``candidate vertices"}. From these we select another grid vertex which will be the start point for the next cylinder.
%\begin{figure}
%\centering
% \includegraphics[width=0.25\textwidth]{imagesMetaTracts/angle-diff.png}
%\caption{Angle between the local orientation of two cylinders that are part of the same MetaTract.}
%\label{fig:angle-diff}
%\end{figure}
The order of the candidate vertices is based on the following characteristics:

@@

\begin{figure}[tb]
\centering
	%\includegraphics[width=0.45\textwidth]{imagesMT2014/crop-16/length_dist.png}
		\includegraphics[width=0.45\textwidth,  trim = 0mm 100mm 0mm 50mm, clip]{imagesMT2014/image_length}
	\caption{(A) MetaTracts colored according to mean orientations. (B) length distribution of individual MetaTracts for a particular bundle.(unit for length is grid cube edge length). A single fiber bundle is made of MetaTracts of varying length}
	\label{fig:length_distribution}
\end{figure}

@@

\begin{figure}[tb] 
  \centering  	
 %\includegraphics[width=0.45\textwidth]{imagesMT2014/crop-16/orientation_cluster.PNG}
 \includegraphics[width=0.5\textwidth]{imagesMT2014/image_clustering.png}
  	\caption{(B) Result of K means clustering with data points projected to the top three eigenvectors as major axes. (A) MetaTracts belonging to orientation cluster 1. (C) MetaTracts belonging to orientation cluster 2. MetaTracts in gray(A,C) show context.
  	(D,E) shows the result of distance based clustering on the orientation cluster. (F) shows the combined result. }
  \label{fig:orientation_clustering}
  \end{figure}

@@

The directed Hausdorff distance is very sensitive to outliers in the data.
However, because MetaTracts after orientation clustering are constructed using cylinders with similar orientations, they are not plagued by outliers.
To cluster based on MetaTract proximity, we used single linkage hierarchical clustering.
Hierarchical clustering has a single parameter $h$, the desired number of clusters.
In hierarchical clustering each objects starts in its own cluster and clusters are merged based on some criterion.
Clusters are merged until there are only $h$ clusters left.
Hierarchical clustering is intuitive since it is easy to trace how clusters are formed and merged.
Single linkage clustering finds pairs of objects $p \in P$ and $q \in Q$ where $P \neq Q$ which are closer than other such pairs, and merges the containing clusters $P$ and $Q$.
We found that single linkage hierarchical clustering had two major drawbacks.
First, the clustering would produce some small clusters of just a few MetaTracts.
These MetaTracts were anomalies caused by noise in the data and did not represent true fiber bundles.
Second, if two fiber bundles were parallel for some of their length and then separated, they would sometimes be clustered into the same fiber bundle.
A short MetaTract which was parallel to both and did not extend into the separation region could form a link between the two fiber bundles, causing them to be clustered into a single bundle.
To address the problem of small clusters, we applied hierarchical clustering and then identified small clusters with few MetaTracts.
We removed the MetaTracts that were in those clusters from the data set and reapplied hierarchical clustering.
To address the problem of short MetaTracts joining different fiber bundles, we applied hierarchical clustering and then removed the
shortest fibers (fibers less than 0.6 times the median length) in each bundle. We then reapplied hierarchical clustering.
We repeated both until a steady state of clusters is reached and no new small fibers can be removed. The results of clustering the orientation clusters Fig.~\ref{fig:orientation_clustering}(A,C) are shown in Fig.~\ref{fig:orientation_clustering}(D,E) respectively. Note how the purple cluster Fig.~\ref{fig:orientation_clustering}(E) bundle ``7'' is extracted well.
\begin{figure}
\centering
		%\includegraphics[width=0.4\textwidth]{imagesMT2014/garb1}~
		\includegraphics[width=0.5\textwidth, trim = 0mm 70mm 0mm 20mm, clip]{imagesMT2014/image7b.pdf}
	%\includegraphics[width=0.45\textwidth]{imagesMT2014/crop-16/bundles}~ 
	%\includegraphics[height=0.35\textwidth]{imagesMT2014/crop-16/vol_mesh.png} 
	\caption{(A) Volume rendering of data set Figure~\ref{fig:data-char}A.(B) shows a single slice of the data set. (C) shows two of the extracted meshes together. (D) shows the meshes separately.}
	\label{fig:crop-16-decomp}
\end{figure}

@@

\begin{figure}[t]
\centering
	\includegraphics[width=0.45\textwidth]{imagesMT2014/image8b}
	\caption{Data set with flat thin and compact bundles. (A) shows the volume rendering and a 2D slice with one of the boundaries marked in green, (B) shows the clusters according to individual orientation. (C) shows the complete result. (D) shows the voxelization of (C).}
	\label{fig:prepreg}
\end{figure}

>>>>

\begin{figure}[tb]
%\includegraphics[width=0.45\textwidth]{imagesMT2014/algo.png}
%\includegraphics[width=0.5\textwidth,  trim = 0mm 300mm 0mm 0mm, clip,]{imagesMT2014/image_algo}
\includegraphics[width=0.5\textwidth]{images_pvis/workflow}
\caption{Flow chart of the MetaTracts approach for fiber bundle extraction}
\label{fig:flowchart}
\end{figure}

@@

\subsection {MetaTracts generation}
\label{subsec:MetaTracts-generation}
In $\mathbb{R}^2$ all the above properties hold except that cylinders are replaced by rectangles. We explain the process of MetaTracts generation in $\mathbb{R}^2$. The procedure trivially extends to $\mathbb{R}^3$. The reliable Hessian step  generates a local orientation in $\mathbb{R}^2$ and a $R_H$[0,1]. In Figure \ref{fig:algo} all regions which have reliable Hessian less than the threshold have been marked in blue(Fig.~\ref{fig:algo}B). Let the seed point associated with the MetaTract be grid point $C_{p}$ (Property 2, Section~\ref{subsec:mtprop}). The local orientation at $C_p$ as computed in the step above is $N_P$ and is given by the dark green arrow (Fig.~\ref{fig:algo}C). We generate a rectangle of length L and radius R, the green dashed rectangle shows the rectangle generated at $C_p$ (Property 3 and 5, Section~\ref{subsec:mtprop}). 
\begin{figure}[tb]
  \begin{center}
 \includegraphics[ width=0.4\textwidth]{imagesMT2014/image4}
  \end{center}
  \vspace{-1em}
 \caption{MetaTracts generation: (B) MetaTract generation, region with $R_H$ less than threshold. (A) close up of marked region in (B). (C) Parts of a single MetaTract.}
  \label{fig:algo}
    \vspace{-1.5em}
\end{figure}
The set of vertices which are in the green region but not overlapped by blue are possible candidates for the start point of the next cylinder in a particular MetaTract. We call these vertices \textit{``candidate vertices"}. From these we select another grid vertex which will be the start point for the next cylinder.
%\begin{figure}
%\centering
% \includegraphics[width=0.25\textwidth]{imagesMetaTracts/angle-diff.png}
%\caption{Angle between the local orientation of two cylinders that are part of the same MetaTract.}
%\label{fig:angle-diff}
%\end{figure}
The order of the candidate vertices is based on the following characteristics:

@@

\begin{figure}[tb]
\centering
	\includegraphics[width=0.45\textwidth]{images_pvis/figure5.png}
	%\includegraphics[width=0.45\textwidth,  trim = 0mm 100mm 0mm 50mm, clip]{imagesMT2014/image_length}
	\caption{(A) MetaTracts colored according to mean orientations. (B) length distribution of individual MetaTracts for a particular bundle.(unit for length is grid cube edge length). A single fiber bundle is made of MetaTracts of varying length}
	\label{fig:length_distribution}
\end{figure}

@@

\begin{figure}[tb] 
  \centering  	
  %\includegraphics[width=0.45\textwidth]{images_pvis/figure6.png}
% \includegraphics[width=0.5\textwidth]{imagesMT2014/image_clustering.png}
  	\caption{(B) Result of K means clustering with data points projected to the top three eigenvectors as major axes. (A) MetaTracts belonging to orientation cluster 1. (C) MetaTracts belonging to orientation cluster 2. MetaTracts in gray(A,C) show context.
  	(D,E) shows the result of distance based clustering on the orientation cluster. (F) shows the combined result. }
  \label{fig:orientation_clustering}
  \end{figure}

@@

The directed Hausdorff distance is very sensitive to outliers in the data.
However, because MetaTracts after orientation clustering are constructed using cylinders with similar orientations, they are not plagued by outliers.
To cluster based on MetaTract proximity, we used single linkage hierarchical clustering.
Hierarchical clustering has a single parameter $h$, the desired number of clusters.
In hierarchical clustering each objects starts in its own cluster and clusters are merged based on some criterion.
Clusters are merged until there are only $h$ clusters left.
Hierarchical clustering is intuitive since it is easy to trace how clusters are formed and merged.
Single linkage clustering finds pairs of objects $p \in P$ and $q \in Q$ where $P \neq Q$ which are closer than other such pairs, and merges the containing clusters $P$ and $Q$.
We found that single linkage hierarchical clustering had two major drawbacks.
First, the clustering would produce some small clusters of just a few MetaTracts.
These MetaTracts were anomalies caused by noise in the data and did not represent true fiber bundles.
Second, if two fiber bundles were parallel for some of their length and then separated, they would sometimes be clustered into the same fiber bundle.
A short MetaTract which was parallel to both and did not extend into the separation region could form a link between the two fiber bundles, causing them to be clustered into a single bundle.
To address the problem of small clusters, we applied hierarchical clustering and then identified small clusters with few MetaTracts.
We removed the MetaTracts that were in those clusters from the data set and reapplied hierarchical clustering.
To address the problem of short MetaTracts joining different fiber bundles, we applied hierarchical clustering and then removed the
shortest fibers (fibers less than 0.6 times the median length) in each bundle. We then reapplied hierarchical clustering.
We repeated both until a steady state of clusters is reached and no new small fibers can be removed. The results of clustering the orientation clusters Fig.~\ref{fig:orientation_clustering}(A,C) are shown in Fig.~\ref{fig:orientation_clustering}(D,E) respectively. Note how the purple cluster Fig.~\ref{fig:orientation_clustering}(E) bundle ``7'' is extracted well.
\begin{figure}
\centering
		\includegraphics[width=0.4\textwidth]{images_pvis/figure7}
		%\includegraphics[width=0.5\textwidth, trim = 0mm 70mm 0mm 20mm, clip]{imagesMT2014/image7b.pdf}
	%\includegraphics[width=0.45\textwidth]{imagesMT2014/crop-16/bundles}~ 
	%\includegraphics[height=0.35\textwidth]{imagesMT2014/crop-16/vol_mesh.png} 
	\caption{(A) Volume rendering of data set Figure~\ref{fig:data-char}A.(B) shows a single slice of the data set. (C) shows two of the extracted meshes together. (D) shows the meshes separately.}
	\label{fig:crop-16-decomp}
\end{figure}

@@

\begin{figure}[t]
\centering
	%\includegraphics[width=0.45\textwidth]{imagesMT2014/image8b}
	\includegraphics[width=0.4\textwidth, trim = 0mm 30mm 0mm 0mm, clip,]{images_pvis/figure8}
	\caption{Data set with flat thin and compact bundles. (A) shows the volume rendering and a 2D slice with one of the boundaries marked in green, (B) shows the clusters according to individual orientation. (C) shows the complete result. (D) shows the voxelization of (C).}
	\label{fig:prepreg}
\end{figure}
